
```{r}
names(gene)[1] <- names(ct)[1]
full_df = merge(ct, gene, by = 'subject_id')
```

```{r}
colnames(full_df[, 152:157])
```


```{r}
full_df$apoe <- as.factor(full_df$apoe)
levels(full_df$apoe)

full_df$bin1 <- as.factor(full_df$bin1)
levels(full_df$bin1)

full_df$clu <- as.factor(full_df$clu)
levels(full_df$clu)

full_df$abca7 <- as.factor(full_df$abca7)
levels(full_df$abca7)

full_df$cr1 <- as.factor(full_df$cr1)
levels(full_df$cr1)

full_df$picalm <- as.factor(full_df$picalm)
levels(full_df$picalm)
```

levels are in correct order

```{r}
full_df <- full_df[complete.cases(full_df),]
```


```{r}
# Compute the analysis of variance
res.aov <- aov(left_G_and_S_frontomargin ~ apoe, data = full_df)
# Summary of the analysis
summary(res.aov)
```


As the p-value is less than the significance level 0.05, we can conclude that there are significant differences between the groups highlighted with “*" in the model summary.

In one-way ANOVA test, a significant p-value indicates that some of the group means are different, but we don’t know which pairs of groups are different.

It’s possible to perform multiple pairwise-comparison, to determine if the mean difference between specific pairs of group are statistically significant.


ALL brain areas vs all genes
```{r}
brain_area <- names(full_df[, 2:149])
genotype <- names(full_df[, 152:157])
genotype[7] <- 'sex'
```


```{r}
library("broom")
```

#one way anove
1. all brain areas vs single genotype 

```{r}
aov_single <- as.data.frame(matrix(data = NA, nrow = length(brain_area), ncol = length(genotype)+1))
names(aov_single)[1] <- 'brain_area'
names(aov_single)[2] <- genotype[1]
names(aov_single)[3] <- genotype[2]
names(aov_single)[4] <- genotype[3]
names(aov_single)[5] <- genotype[4]
names(aov_single)[6] <- genotype[5]
names(aov_single)[7] <- genotype[6]
names(aov_single)[8] <- genotype[7]
aov_single$brain_area <- brain_area
```

```{r}
for (g in 1:length(genotype)){
  for (br in 1:length(brain_area)){
    res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[g]]])
    s <- tidy(res.aov)
    aov_single[br, (g+1)] <- s$p.value[1]
  }
}

aov_single$bonferroni_apoe <- p.adjust(aov_single$apoe, method = "bonferroni")
aov_single$fdr_apoe <- p.adjust(aov_single$apoe, method = "fdr")

aov_single$bonferroni_bin1 <- p.adjust(aov_single$bin1, method = "bonferroni")
aov_single$fdr_bin1 <- p.adjust(aov_single$bin1, method = "fdr")

aov_single$bonferroni_clu <- p.adjust(aov_single$clu, method = "bonferroni")
aov_single$fdr_clu <- p.adjust(aov_single$clu, method = "fdr")

aov_single$bonferroni_abca7 <- p.adjust(aov_single$abca7, method = "bonferroni")
aov_single$fdr_abca7 <- p.adjust(aov_single$abca7, method = "fdr")

aov_single$bonferroni_cr1 <- p.adjust(aov_single$cr1, method = "bonferroni")
aov_single$fdr_cr1 <- p.adjust(aov_single$cr1, method = "fdr")

aov_single$bonferroni_picalm <- p.adjust(aov_single$picalm, method = "bonferroni")
aov_single$fdr_picalm <- p.adjust(aov_single$picalm, method = "fdr")

aov_single$bonferroni_sex <- p.adjust(aov_single$sex, method = "bonferroni")
aov_single$fdr_sex <- p.adjust(aov_single$sex, method = "fdr")

aov_single[aov_single$fdr_sex < 0.05, c(1, 22)]
```

#two way anova
Two-way ANOVA test is used to evaluate simultaneously the effect of two grouping variables (A and B) on a response variable.
2. all brain areas vs double genotypes

one dependent variable - cortical thickness
and 2 independent variables - 2 independent categorical variable

2 independent categorical variable in additive model => brain area ~ gene1 + gene2
in the situation where the interaction is not significant you should use the additive model.

2 categorical variables interact to create an synergistic effect => brain area ~ gene1 *+* gene2

```{r}
      res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[1]]])
      tidy(res.aov)


      res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[2]]])
      tidy(res.aov)

      res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[1]]]+full_df[[genotype[2]]])
      tidy(res.aov)
      
        
      res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[1]]]*full_df[[genotype[2]]])
      tidy(res.aov)

```




```{r}
aov_double <- as.data.frame(matrix(data = NA, nrow = length(brain_area), ncol = 22)) #ncol = C_2(7)
names(aov_double)[1] <- 'brain_area'
aov_double$brain_area <- brain_area

k <- 2
for (i in 1:(length(genotype)-1)){
  for (j in (i+1):length(genotype)){
        names(aov_double)[k] <- paste(genotype[i], genotype[j])
        k <- k + 1
  }
}

```


```{r}
k <- 2
for (i in 1:(length(genotype)-1)){
  for (j in (i+1):length(genotype)){
    for (br in 1:length(brain_area)){
      res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[i]]]*full_df[[genotype[j]]])
      s <- tidy(res.aov)
      #aov_double[br, k] <- paste(s$p.value[1], s$p.value[2], s$p.value[3])
      aov_double[br, k] <- s$p.value[3]
    }
    k <- k + 1
  }
}

for (col in 23:43){
  aov_double[, col] <- p.adjust(aov_double[, (col-21)], method = "fdr")
}

aov_double[aov_double$V23 < 0.05, c(1, 2, 23)]

aov_double[aov_double$V24 < 0.05, c(1, 3, 24)]

aov_double[aov_double$V26 < 0.05, c(1, 5, 26)]

aov_double[aov_double$V27 < 0.05, c(1, 6, 27)]

aov_double[aov_double$V34 < 0.05, c(1, 13, 34)]


```

```{r}
boxplot(right_G_and_S_frontomargin ~ apoe * bin1, data=full_df, frame = FALSE, 
        col = c("#00AFBB", "#E7B800"), ylab="cortical thickness")
```

```{r}
      res.aov <- aov(full_df[[brain_area[75]]] ~ full_df[[genotype[1]]])
      tidy(res.aov)


      res.aov <- aov(full_df[[brain_area[75]]] ~ full_df[[genotype[2]]])
      tidy(res.aov)

      res.aov <- aov(full_df[[brain_area[75]]] ~ full_df[[genotype[1]]]+full_df[[genotype[2]]])
      tidy(res.aov)
      
        
      res.aov <- aov(full_df[[brain_area[75]]] ~ full_df[[genotype[1]]]*full_df[[genotype[2]]])
      tidy(res.aov)

```
It can be seen that the two main effects are close to be statistically significant, as well as their interaction.


3. all brain areas vs triple genotypes
#Three-way anova 
```{r}
for (n in 1:(length(genotype)-2)){
  for (i in (n+1):(length(genotype)-1)){
    for (j in (i+1):length(genotype)){
      for (br in 1:length(brain_area)){
        res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[n]]]*full_df[[genotype[i]]]*full_df[[genotype[j]]])
        s <- tidy(res.aov)
        if (mean(s$p.value, na.rm = TRUE) < 0.1){
          print(s)
          print(paste(br, genotype[n], genotype[i], genotype[j]))
          }
      }
    }
  }
}

```

```{r}
three_way_anova <- full_df[, c(1, 145, 152, 153, 155)]
three_way_anova <- three_way_anova %>%  unite("gene_combo", apoe:abca7, remove = FALSE)
```


```{r}
ggboxplot(three_way_anova, x = "gene_combo", y = "right_S_suborbital", 
          color = "gene_combo", 
          ylab = "right_S_suborbital", xlab = "apoe_bin1_abca7") + rotate_x_text(45)
```
if I change to (mean(s$p.value, na.rm = TRUE) < 0.2) there is around 20 results

```{r}
for (n in 1:(length(genotype)-2)){
  for (i in (n+1):(length(genotype)-1)){
    for (j in (i+1):length(genotype)){
      for (br in 1:length(brain_area)){
        res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[n]]]*full_df[[genotype[i]]]*full_df[[genotype[j]]])
        s <- tidy(res.aov)
        if (s$p.value[7] < 0.05){
          print(s)
          print(paste(br, genotype[n], genotype[i], genotype[j]))
          }
      }
    }
  }
}

```

4. all brain aress vs quadruple genotypes


```{r}
for (k in 1:(length(genotype)-3)){
  for (n in (k+1):(length(genotype)-2)){
    for (i in (n+1):(length(genotype)-1)){
      for (j in (i+1):length(genotype)){
        for (br in 1:length(brain_area)){
          res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[k]]] * full_df[[genotype[n]]] * full_df[[genotype[i]]] * 
                           full_df[[genotype[j]]])
          s <- tidy(res.aov)
          if (mean(s$p.value, na.rm = TRUE) < 0.15){
            print(s)
            #print(br, genotype[k], genotype[n], genotype[i], genotype[j])}
          }
        }
      }
    }
  }
}
#NO RESULT
```


5. all brain areas vs 5 genetopypes

```{r}
for (l in 1:(length(genotype)-4)){
  for (k in (l+1):(length(genotype)-3)){
    for (n in (k+1):(length(genotype)-2)){
      for (i in (n+1):(length(genotype)-1)){
        for (j in (i+1):length(genotype)){
          for (br in 1:length(brain_area)){
            res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[l]]] * full_df[[genotype[k]]] * full_df[[genotype[n]]] *
                             full_df[[genotype[i]]] * full_df[[genotype[j]]])
            s <- tidy(res.aov)
            if (mean(s$p.value, na.rm = TRUE) < 0.15){
              print(s)
              #print(br, genotype[k], genotype[n], genotype[i], genotype[j])}
            }
          }
        }
      }
    }
  }
}
#NO RESULT
```

6. all brain areas vs 6 genotypes

```{r}
for (m in 1:(length(genotype)-5)){
  for (l in (m+1):(length(genotype)-4)){
    for (k in (l+1):(length(genotype)-3)){
      for (n in (k+1):(length(genotype)-2)){
        for (i in (n+1):(length(genotype)-1)){
          for (j in (i+1):length(genotype)){
            for (br in 1:length(brain_area)){
              res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[m]]] * full_df[[genotype[l]]] * full_df[[genotype[k]]] *
                               full_df[[genotype[n]]] * full_df[[genotype[i]]] * full_df[[genotype[j]]])
              s <- tidy(res.aov)
              if (mean(s$p.value, na.rm = TRUE) < 0.1){
                print(s)
                #print(br, genotype[k], genotype[n], genotype[i], genotype[j])}
              }
            }
          }
        }
      }
    }
  }
}
#NO RESULT
```

7. all brain areas vs 7 genotypes

```{r}
for (br in 1:length(brain_area)){
  res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[1]]] + full_df[[genotype[2]]]
                 + full_df[[genotype[3]]] + full_df[[genotype[4]]] + full_df[[genotype[5]]] + full_df[[genotype[6]]] + full_df[[genotype[7]]])
  s <- tidy(res.aov)
  if (mean(s$p.value, na.rm = TRUE) < 0.1){
    print(s)
  }
}

#NO RESULT

for (br in 1:length(brain_area)){
  res.aov <- aov(full_df[[brain_area[br]]] ~ full_df[[genotype[1]]] * full_df[[genotype[2]]]
                 * full_df[[genotype[3]]] * full_df[[genotype[4]]] * full_df[[genotype[5]]] * full_df[[genotype[6]]] * full_df[[genotype[7]]])
  s <- tidy(res.aov)
  if (mean(s$p.value, na.rm = TRUE) < 0.1){
    print(s)
  }
}
#NO RESULT

```




AVG CORTICAL THICKNESS 

NO SIGNIFICANT RELATIONSGIP WITH GENOTYPES WERE REVEALED

```{r}
names(full_df)[75]
```


```{r}
for (g in 1:length(genotype)){ #[2:75] left or [76;149] right or [2:149] all
    res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~ full_df[[genotype[g]]])
    s <- tidy(res.aov)
    if (s$p.value[1] < 0.05){
      print(s)
    }
}
```

No dependency between avg cortical thickness and any single genotype

```{r}
for (i in 1:(length(genotype)-1)){
  for (j in (i+1):length(genotype)){
      res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~ full_df[[genotype[i]]]*full_df[[genotype[j]]])
      s <- tidy(res.aov)
      if (s$p.value[3] < 0.05){
      print(s)
      print(paste(genotype[i], genotype[j]))
      }
  }
}
```

```{r}
for (n in 1:(length(genotype)-2)){
  for (i in (n+1):(length(genotype)-1)){
    for (j in (i+1):length(genotype)){
        res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~
                         full_df[[genotype[n]]]*full_df[[genotype[i]]]*full_df[[genotype[j]]])
        s <- tidy(res.aov)
        if (mean(s$p.value, na.rm = TRUE) < 0.1){
          print(s)
          print(paste(br, genotype[n], genotype[i], genotype[j]))
          }
    }
  }
}

```


```{r}
for (k in 1:(length(genotype)-3)){
  for (n in (k+1):(length(genotype)-2)){
    for (i in (n+1):(length(genotype)-1)){
      for (j in (i+1):length(genotype)){
        res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~
                         full_df[[genotype[k]]] * full_df[[genotype[n]]] * full_df[[genotype[i]]] * 
                           full_df[[genotype[j]]])
          s <- tidy(res.aov)
          if (mean(s$p.value, na.rm = TRUE) < 0.1){
            print(s)
            #print(br, genotype[k], genotype[n], genotype[i], genotype[j])}
          
        }
      }
    }
  }
}

```



```{r}
for (l in 1:(length(genotype)-4)){
  for (k in (l+1):(length(genotype)-3)){
    for (n in (k+1):(length(genotype)-2)){
      for (i in (n+1):(length(genotype)-1)){
        for (j in (i+1):length(genotype)){
          res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~
                           full_df[[genotype[l]]] * full_df[[genotype[k]]] * full_df[[genotype[n]]] *
                             full_df[[genotype[i]]] * full_df[[genotype[j]]])
            s <- tidy(res.aov)
            if (mean(s$p.value, na.rm = TRUE) < 0.15){
              print(s)
              #print(br, genotype[k], genotype[n], genotype[i], genotype[j])}
           
          }
        }
      }
    }
  }
}

```


```{r}
for (m in 1:(length(genotype)-5)){
  for (l in (m+1):(length(genotype)-4)){
    for (k in (l+1):(length(genotype)-3)){
      for (n in (k+1):(length(genotype)-2)){
        for (i in (n+1):(length(genotype)-1)){
          for (j in (i+1):length(genotype)){
            res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~
                             full_df[[genotype[m]]] * full_df[[genotype[l]]] * full_df[[genotype[k]]] *
                               full_df[[genotype[n]]] * full_df[[genotype[i]]] * full_df[[genotype[j]]])
              s <- tidy(res.aov)
              if (mean(s$p.value, na.rm = TRUE) < 0.1){
                print(s)
                #print(br, genotype[k], genotype[n], genotype[i], genotype[j])}
              
            }
          }
        }
      }
    }
  }
}

```


```{r}
res.aov <- aov(rowMeans(full_df[, 2:149], na.rm = TRUE) ~ full_df[[genotype[1]]] + full_df[[genotype[2]]]
                 + full_df[[genotype[3]]] + full_df[[genotype[4]]] + full_df[[genotype[5]]] + full_df[[genotype[6]]] + full_df[[genotype[7]]])
  s <- tidy(res.aov)
  if (mean(s$p.value, na.rm = TRUE) < 0.1){
    print(s)
  }

```







