
#AN EXAMPLE THAT ANOVA WAS REALLY USEFUL PROCEDURE

```{r}
model2 <- lm(rowMeans(ct[, -1], na.rm = TRUE, dims = 1) ~ gene$apoe + gene$bin1 + gene$abca7)
summary(model2)
```

right_S_suborbital", xlab = "apoe_bin1_abca7"

```{r}
model2 <- lm(ct$right_S_suborbital ~ gene$apoe + gene$bin1 + gene$abca7)
summary(model2)
```

So we can notice that ANOVA test was useful since there was no relashinsip discovered for mentioned set of genes and avg cortical thickness but there is significant p-value for right_S_suborbital brain area and the mentioned set of genes


#AVG cortical thickness

[1] "apoe clu"

```{r}
model2 <- lm(rowMeans(ct[, -1], na.rm = TRUE, dims = 1) ~ gene$apoe + gene$clu)
summary(model2)
```

[1] "apoe picalm"
```{r}
model2 <- lm(rowMeans(ct[, -1], na.rm = TRUE, dims = 1) ~ gene$apoe + gene$picalm)
summary(model2)
```

NO LR with good result!



#SINGLE BRAIN AREA 
ONE_WAY ANOVA RESULTS
BRAIN AREA AND SEX

left_G_and_S_cingul.Mid.Ant	0.0092
```{r}
model2 <- lm(ct$left_G_and_S_cingul.Mid.Ant ~ gene$sex)
summary(model2)
```

left_S_collat_transv_post	0.0219
left_S_oc_middle_and_Lunatus	0.0379
left_S_oc.temp_med_and_Lingual	0.0380
left_S_temporal_transverse	0.0219
right_G_and_S_cingul.Mid.Post	0.0379
right_S_oc.temp_med_and_Lingual	0.0002
```{r}
ggboxplot(full_df, x = "sex", y = "right_S_oc.temp_med_and_Lingual", 
          color = "sex", 
          ylab = "right_S_oc.temp_med_and_Lingual", xlab = "sex") 
```
right_S_temporal_transverse	0.0219



TWO_WAY ANOVA RESULT 
apoe+bin1	
brain_area	adjusted p-value
left_G_subcallosal	0.0412
right_G_and_S_frontomargin	0.0035

```{r}
model2 <- lm(ct$left_G_subcallosal ~ gene$apoe)
summary(model2)
```

```{r}
model2 <- lm(ct$left_G_subcallosal ~ gene$bin1)
summary(model2)
```

parameters for apoe and parameters for bin1. But no interaction.
```{r}
model2 <- lm(ct$left_G_subcallosal ~ gene$apoe + gene$bin1)
summary(model2)
```

```{r}
grep("bin1", colnames(full_df))
```

```{r}
apoebin1_plot <- full_df[, c(33, 152, 153)]
apoebin1_plot$concat <- paste(as.character(apoebin1_plot$apoe), as.character(apoebin1_plot$bin1))
```

```{r}
ggboxplot(apoebin1_plot, x = "concat", y = "left_G_subcallosal", 
          color = "concat", 
          ylab = "left_G_subcallosal", xlab = "apoe_bin1") + rotate_x_text(45)
```


include these variables and the interactions between them
```{r}
model2 <- lm(ct$left_G_subcallosal ~ gene$apoe * gene$bin1)
summary(model2)
```
You need to know why some parameters will appear to be 'missing' from the output of the call to lm. This is due to the way factors enter the design matrix;


Only the interactions

```{r}
model2 <- lm(ct$left_G_subcallosal ~ gene$apoe : gene$bin1)
summary(model2)
```

some parameters will appear to be 'missing' from the output of the call to lm. This is due to the way factors enter the design matrix


USAGE OF ANOVA TEST TO PROVE THAT INCLUDING ADDITIONAL VARIABLE TO MODEL 

http://strata.uga.edu/8370/lecturenotes/multipleRegression.html

 the simpler model always contains a subset of the parameters of the more complicated model
 
 The idea is that the difference between the two models has a quantity called the extra sum of squares, equal to the reduction in the unexplained sum of squares produced by the additional model terms. This can be converted to a variance, which can be scaled against the unexplained variance term for the full model. A ratio of variances can be tested with an F-test; hence this problem can be solved with an ANOVA. Testing for the significance of additional variables between two models:

```{r}
model1 <- lm(ct$left_G_subcallosal ~ gene$apoe)
model2 <- lm(ct$left_G_subcallosal ~ gene$bin1)
model3 <- lm(ct$left_G_subcallosal ~ gene$apoe : gene$bin1)
anova(model1, model2, model3)
```

```{r}
model2 <- lm(ct$left_G_subcallosal ~ gene$apoe * gene$bin1)
```

```{r}
library("olsrr")
```

Residual QQ Plot
Graph for detecting violation of normality assumption.

```{r}
ols_plot_resid_qq(model2)
```
Residual Normality Test
Test for detecting violation of normality assumption.

```{r}
ols_test_normality(model2)
```

Correlation between observed residuals and expected residuals under normality.

```{r}
ols_test_correlation(model2)
```
Histogram of residuals for detecting violation of normality assumption.

```{r}
ols_plot_resid_hist(model2)

```







